---
title: "Simulating Policy Scenarios using Aggregate Model"
author: "Abhinav Saxena"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example Workflow}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=7,
  fig.height=5
)
```

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(aggregate.model)
library(dplyr)
library(purrr)
library(ggplot2)
library(here)
```

# Simulating Policy Scenarios using Aggregate Model

## Applying the Aggregate Model to Real World Scenarios

While tackling the economic policy aspects of climate change,
policymakers are required to take certain steps or announce policy
measures that aim at reducing emissions from a particular economic
activity. These policy measures can include announcing subsidies, tax
breaks, investments in the activities that promote an
environment-friendly outlook to the activity. On the other hand, the
activities that are carbon-intensive and have a high environmental
impact need to be reduced or completely abandoned as there is no more
capacity for the planet to accommodate more emissions. The relevant
policy measures in this case can be quantity limits, taxing the activity
(or raising the tax), and in some cases even cutting off the activity.

In this vignette, I will simulate some of these policy scenarios using
the Aggregate Model and will try to calculate the trajectory of
emissions and the savings in emissions if these were to be exogenously
pushed by the policymakers.

### Policy Scenario 1 : Metal Production in Germany

Germany is the biggest producer of Metals in the European Union and is
one of the top producers in the world markets too. Metal Production is
also one of the most energy and emissions intensive economic activities.

In this scenario, I will simulate the effect of a policy proposal that
aims at cutting metal production in Germany to 50% of its 2015 values by
2027, as part of its commitments to an international climate agreement.

```{r, Creating a dictionary for Germany}

dict_de <- dict %>%
  mutate(geo = "DE")
```

#### Adding Variables to the Dictionary

Once the dictionary is modified for our desired country, we can start by
adding some variables that may be important for our analysis. In this
case, I have selected the Index of Metal Production, Index of Industrial
Production, and the Index of Plastic Production in Germany where the
2015 production value is set as a reference/index value of 100.

```{r, Adding variables to our dictionary}

dict_de <- dict_de %>%
  bind_rows(tibble(
  model_varname = "Metal_Prod",
  full_name = "Index of Manufacture of basic metals and fabricated metal products",
  database  = "eurostat",
  variable_code = "PROD",
  dataset_id = "sts_inpr_q",
  var_col = "indic_bt",
  freq = "q",
  geo = "DE",
  unit = "I15",
  s_adj = "NSA",
  nace_r2 = "C24_C25"))

dict_de <- dict_de %>%
  bind_rows(tibble(
    model_varname = "Ind_Prod",
    full_name = "An index of Industrial Production",
    database  = "eurostat",
    variable_code = "PROD",
    dataset_id = "sts_inpr_q",
    var_col = "indic_bt",
    freq = "q",
    geo = "DE",
    unit = "I15",
    s_adj = "NSA",
    nace_r2 = "B-D"))

dict_de <- dict_de %>%
  bind_rows(tibble(
    model_varname = "PlaProd",
    full_name = "An index of Plastic Production",
    database  = "eurostat",
    variable_code = "PROD",
    dataset_id = "sts_inpr_q",
    var_col = "indic_bt",
    freq = "q",
    geo = "DE",
    unit = "I15",
    s_adj = "NSA",
    nace_r2 = "C222"))
```

#### Specification

For this particular illustrative scenario, we have used a three equation
specification of the model:

```{r specification}
specification <- dplyr::tibble(
  type = c(
    "n",
    "n",
    "n"
  ),
  dependent = c(
    "Ind_Prod",
    "EmiCO2Industry",
    "EmiCO2Combustion"
  ),
  independent = c(
    "Metal_Prod + PlaProd + Import",
    "HICP_Energy + Metal_Prod + Export",
    "FinConsExpHH + HICP_Gas + EmiCO2Industry + Ind_Prod"
    )
)
specification
```

In this specification, we have modelled Industrial Production,
Industrial Emissions, and Emissions from Combustion. For these different
dependent variables, there are a set of independent variables that have
been used. But in this simulation, we are most concerned with the effect
of Metal_Prod on EmiCO2Industry.

#### Running the Model

We are now ready to run the model. We have created a new folder called
metal where the fetched data (data_metal) would be saved and this
locally saved data would be used to run the model in subsequent
iterations.

```{r run-model}
model <- run_model(specification = specification,
                   dictionary = dict_de,
                   inputdata_directory = here("data-raw/metal/"),
                   primary_source = "local",
                   save_to_disk = here("data-raw/metal/data_metal.xlsx"),
                   saturation.tpval = 0.001,
                   max.ar = 2,
                   max.dl = 2,
                   quiet = FALSE,
                   saturation = "IIS",
                   constrain.to.minimum.sample = FALSE)
```

#### Evaluating the Model

Now, we can have a look at the model summary. As we can see, the
coefficient of Metal Production is statistically significant at a
significance level of 0.001, thereby indicating a strong positive
relationship between Industrial CO2 Emissions and Metal Production in
Germany.

```{r message=FALSE, warning=FALSE}
library(modelsummary)

class(model)
model_list <- lapply(model$module_collection$model, gets::as.lm)
names(model_list) <- model$module_order$dependent
modelsummary::modelsummary(
  model_list,
  coef_omit = "iis|sis",
  gof_omit = "R",
  notes = "Impulse (IIS) and Step Indicators (SIS) are not shown individually but were activated for all models.",
  stars = TRUE
)
```

##### Module Network

We can show the relationship between the different modules using the
`network()` function.

```{r network, fig.width=7, fig.height=5}
network(model)
```

Each node represents a module and the different colours represent
whether the variable is given by a definition/identity, whether it has
been modelled as an endogenous variable depending on other models, and
whether it is an exogenous variable input to the models.

An solid line arrow means that the variable has been retained during
model selection, while a dashed arrow means that the variable has been
found insignificant and thus dropped during model selection.

#### Forecasting (Simulating Policy Scenarios)

Now, this is where we can simulate policy scenarios based on the
forecasted values of variables we are interested in. Firstly, we can use
the forecast_model() function to forecast values of the different
modules using an autoregressive (AR) function and save the results.
Since, in this scenario assumption, we have 15 quarters to model we have
set the parameter n.ahead=15.

```{r, AR forecast}
f1 <- forecast_model(model, exog_fill_method = "AR", n.ahead = 15, plot.forecast = FALSE)
plot(f1)
```

The following code is crucial for this scenario assumption. We have
assumed that the cuts in production will progress over the quarters at a
constant rate. To calculate this rate and therefore the indexed quantity
of metal production, we have used a simple compounding formula which
estimates a rate of 3.76% decline per quarter. In simple terms, the
production of each subsequent quarter will be 3.76% less than the
preceding quarter so that it finally reaches to 50% (of 2015) by the
last quarter of 2027. It can be noted that this is only one way to reach
this goal.

```{r, Calculating Declining Values}
initial_value <- f1$exog_data_nowcast$Metal_Prod[1]
final_value <- 50 #Final Value is the desired policy implementation. That by 2027, the metal production should be 60% of that of the 2015 value.
periods <- 16  

rate <- (final_value / initial_value)^(1 / (periods - 1)) #In this case, this requires cutting the metal production by around 3.76% of the previous quarter

decreasing_values <- initial_value * rate^(0:(periods - 1))
```

Now we can apply these calculated values as exogenously provided values
to forecast another model with these lower values of metal production to
calculate savings in industrial emissions.

```{r, exogenous data forecasting}
exog_data_high_ready <- f1$exog_data_nowcast %>%
  mutate(Metal_Prod = decreasing_values)

f1_high <- forecast_model(model, exog_predictions = exog_data_high_ready, plot.forecast = FALSE)
```

From this forecast, we can easily retain the data of two variables we
are interested in, namely Industrial Emissions and Metal Production. The
grepl_variables parameter in the plot() function matches the string
combination to automatically capture the desired variable without the
need of writing the exact variable name as it is saved. We can use this
data to create a plot to visually present our forecast results of these
variables under the exogenously provided metal production quantities.

```{r, plotting the forecast}
f1_data <- plot(f1, exclude.exogenous = FALSE, grepl_variables = "Industry|Metal", return.data = TRUE)
f1_high_data <- plot(f1_high, exclude.exogenous = FALSE, grepl_variables = "Industry|Metal", return.data = TRUE)

plotting_scenario_data <- f1_data %>%
  mutate(name = "AR Forecast") %>%
  bind_rows(f1_high_data %>%
              mutate(name = NA) %>%
              mutate(name = case_when(na_item == "EmiCO2Industry" & type == "Endogenous Forecast" ~
              "Scenario Values",
              na_item == "Metal_Prod" & type == "Exogenous Forecast" ~ "Scenario Values",
              TRUE ~ name)))

plotting_scenario_data %>%
  ggplot(aes(x = time, y = values, linetype = name, color = type)) +
  geom_line() +
  coord_cartesian(xlim = c(as.Date("2020-01-01"),as.Date("2027-10-01"))) +
  facet_wrap(~na_item, scales = "free") +
  geom_line(linewidth = 1, na.rm = TRUE) +
  facet_wrap(~.data$na_item, scales = "free") +
  labs(x = "Year", y = "Values") +
  scale_y_continuous(labels = scales::label_comma(), limits = c(0, NA)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())+
  labs(title = "Illustrative Example for Germany - Reduction in Metal Production")
```

Finally calculating the savings in emissions. It can be noted a positive
diff indicates the rise in emissions given the scenario values whereas a
negative diff value suggests a fall in emissions under the scenario. In
this case, we observe a very high diff of 106714 units between the AR
forecasted values of metal production and those given exogenously.

```{r, savings calculation}
f1_data %>%
  mutate(name = "AR Forecast") %>%
  bind_rows(f1_high_data %>%
              mutate(name = "Scenario Values")) %>%
  filter(na_item == "EmiCO2Industry" & type == "Endogenous Forecast") %>%
  summarise(emissions = sum(values, na.rm = TRUE), .by = name) %>%
  mutate(total = sum(emissions),
         diff = c(NA,diff(emissions)),
         rel_diff = diff/total)
```

#### Shiny App

We can also get an overview and summary of the whole aggregate model
results in a Shiny app, which can be opened using the`present_model()`
command. The following code snippet is not executed:

```{r, shiny, eval=FALSE}
present_model(model)
```

### Policy Scenario 2 : Energy Subsidies in Austria

In this scenario, a policy proposal that aims to subsidise energy prices
in Austria is considered. This move has been a popular strategy for
governments, during the times of high energy prices, to relieve
households and businesses from high prices. Conversely, it has
significant implications for GHG emissions.

#### Adding Variables to the Dictionary

In this simulation, I have added only one model to our pre-exisiting
dictionary (dict) on Industrial Production.

```{r, Adding variables to our dictionary Model#2}

new_dict <- dict %>%
  bind_rows(tibble(
    model_varname = "IndProd", 
    full_name = "An index of Industrial Production",
    database  = "eurostat",
    variable_code = "PROD", 
    dataset_id = "sts_inpr_q",
    var_col = "indic_bt", 
    freq = "q", 
    geo = "AT",
    unit = "I15", 
    s_adj = "NSA", 
    nace_r2 = "B-D"))
```

#### Specification

For this particular scenario, a three equation specification of the
model has been used:

```{r specification #2}
spec <- dplyr::tibble(
  type = c(
    "n",
    "n",
    "n"
  ),
  dependent = c(
    "EmiCO2Industry",
    "EmiCO2Combustion",
    "FinConsExpHH"
  ),
  independent = c(
    "HICP_Energy + IndProd + Export",
    "FinConsExpHH + HICP_Energy",
    "FinConsExpGov + HICP_Energy"
  )
)
spec
```

In this specification, I have modelled Industrial Emissions, Emissions
from Combustion, and Household Consumption Expenditure. For these
different dependent variables, certain variables such as Energy Prices
(HICP_Energy), Industrial Production Index, Exports, etc. have been
used. But in this simulation, we are most concerned with the effect of
HICP_Energy on Emissions from Combustion.

#### Running the Model

To avoid using saved data for other models, I have created a new folder
titled "Your_Folder" where the data for this particular model can be
saved and used in the further iterations of this model.

```{r run-model2}
model2 <- run_model(specification = spec,
                   dictionary = new_dict,
                   inputdata_directory = here("data-raw/your_folder/"),
                   primary_source = "local",
                   save_to_disk = here("data-raw/your_folder/data_new.xlsx"),
                   saturation.tpval = 0.001,
                   max.ar = 2,
                   max.dl = 2,
                   quiet = FALSE,
                   saturation = NULL,
                   constrain.to.minimum.sample = FALSE)
```

#### Evaluating the Model

Now, we can have a look at the model summary. The negative coefficient
on HICP_Energy in the EmiCO2Combustion represents a negative
relationship between prices and combustion emissions. This result
follows the simple logical mechanism of higher prices, lower demand and
thus lower use of energy, suggesting lower emissions derived from fossil
based energy combustion.

```{r , mod sum#2, message=FALSE, warning=FALSE}
library(modelsummary)

model_list2 <- lapply(model2$module_collection$model, gets::as.lm)
names(model_list2) <- model2$module_order$dependent
modelsummary::modelsummary(
  model_list2,
  gof_omit = "R",
  title = "Models run for each sub-module for this illustrative example of Austria.",
  stars = TRUE
)
```

##### Module Network

We can show the relationship between the different modules using the
`network()` function.

```{r network2, fig.width=7, fig.height=5}
network(model2)
```

Each node represents a module and the different colours represent
whether the variable is given by a definition/identity, whether it has
been modelled as an endogenous variable depending on other models, and
whether it is an exogenous variable input to the models.

An solid line arrow means that the variable has been retained during
model selection, while a dashed arrow means that the variable has been
found insignificant and thus dropped during model selection.

#### Forecasting (Simulating Policy Scenarios)

In this scenario assumption, we are forecasting the energy prices for 10
quarters and would apply a 15% price cut owing to subsidies as compared
to the forecasted values.

```{r, AR forecast #2}
f2 <- forecast_model(model2, exog_fill_method = "AR", plot.forecast = FALSE)
```

Now we can apply these calculated values as exogenously provided values
to forecast another model with lower energy prices.

```{r, forecasting with exogenous values #2}
exog_data_low <- f2$exog_data_nowcast %>%
  mutate(HICP_Energy = 0.85*HICP_Energy)
f2_low <- forecast_model(model2, exog_predictions = exog_data_low, plot.forecast = FALSE)
```

Now we will retain the data for our two desired variables, i.e.,
Emissions from Combustion and HICP_Energy. We can visualise the plot for
the two scenarios as well.

```{r, forecast plots}
f2_data <- plot(f2, exclude.exogenous = FALSE, grepl_variables = "Combustion|Energy", return.data = TRUE)

f2_low_data <- plot(f2_low, exclude.exogenous = FALSE, grepl_variables = "Combustion|Energy", return.data = TRUE)

plotting_scenario_data <- f2_data %>%
  mutate(name = "AR Forecast") %>%
  bind_rows(f2_low_data %>%
              mutate(name = NA) %>%
              mutate(name = case_when(na_item == "EmiCO2Combustion" & type == "Endogenous Forecast" ~ "Scenario Values",
                                      na_item == "HICP_Energy" & type == "Exogenous Forecast" ~ "Scenario Values",
                                      TRUE ~ name)))
plotting_scenario_data %>%
  ggplot(aes(x = time, y = values, linetype = name, color = type)) +
  geom_line() +
  coord_cartesian(xlim = c(as.Date("2020-01-01"),as.Date("2026-07-01"))) +
  facet_wrap(~na_item, scales = "free") +
  geom_line(linewidth = 1, na.rm = TRUE) +
  facet_wrap(~.data$na_item, scales = "free") +
  labs(x = "Year", y = "Values") +
  scale_y_continuous(labels = scales::label_comma(), limits = c(0, NA)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())
```

Finally calculating the extra emissions that result from this policy. It
can be noted a positive diff, in this case, indicates the rise in
emissions. In this case, we observe a diff of 3118 units between the AR
forecasted values and those given exogenously. Thus, in this particular
model, if a energy subsidy amounting to 15% of the forecasted energy
prices is provided, it would imply extra combustion emissions of 3118
units.

```{r, extra emissions calculation}
f2_data %>%
  mutate(name = "AR Forecast") %>%
  bind_rows(f2_low_data %>%
              mutate(name = "Scenario Values")) %>%
  filter(na_item == "EmiCO2Combustion" & type == "Endogenous Forecast") %>%
  summarise(emissions = sum(values, na.rm = TRUE), .by = name) %>%
  mutate(total = sum(emissions),
         diff = c(NA,diff(emissions)),
         rel_diff = diff/total)
```
